CODECHECK: An open-science initiative for the independent execution of computations underlying research articles.

### 2021-01-11 codecheck-paper.tex
update table
add 024 and 025, update all description texts, update references

We have a curated list of 22 certificates available at
\url{https://codecheck.org.uk/register}. These are a mixture of
reproductions of papers for a variety of reasons and fall into three themes:
(1) `classic' papers from computational neuroscience
(2) COVID-19 modelling preprints, and
(3) GIScience.  
The first theme was the initial set of papers used to explore the concept
of CODECHECK. The idea was to take very well known historical articles from a 
specific domain, in this case the work area of co-author SJE, possibly 
as as a conversation starter for potential collaborating journals.
The second theme is rather opportunistic. As the COVID-19 pandemic swept
over the whole Earth, we hoped to contribute a small part to science's answer
to the unknown virus. The checks were solicited through community interaction
%Several COVID-19 models have been done during the time of peer review of these papers,
and under our initiative rather than requested from journal, but the certificate
has then been acknowledged/cited in the paper \cite{Davies2020-vj}.
The third theme represents co-author DN's service as Reproducibility
Reviewer at the AGILE conference series.
Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
established a process for reproducing workflows at the AGILE conference
series \cite{nust_improving_2020}.
While using slightly different terms and infrastructure, e.g., ``reproducibility
reports'' are published on OSF instead of certificates on Zenodo, the 
reproducibility reviews adhere to the codecheck principles.
Furthermore, a small number of checks was completed as part of a peer reviews
for a GIScience journals.

@@

\begin{table}
  \centering

  \begin{tabular}{llp{11cm}}
    \toprule
    Certificate & Research area & Description \\ \midrule
    2020-001 \cite{cert-2020-001} &
    Machine learning & CODECHECK undertaken post acceptance of
    manuscript and before its publication in \textit{Gigascience}
    \cite{Piccolo2020-lo}. \\
    2020-002  \cite{cert-2020-002} & neuroscience & Paper from 1997
 showing unsupervised learning from natural images \cite{Hancock1992-mp}.
 Code written for this project. \\
    2020-003  \cite{cert-2020-003} & neuroscience &  Classic paper
                                                     from 1982 on
                                                     models of
                                                     associative
                                                     memory
                                                     \cite{Hopfield1982-mz}.
    (Simulation code was written specifically for this project.)\\
    2020-004  \cite{cert-2020-004} & neuroscience & Cart-pole balancing problem solved using reinforcement learning \cite{Barto1983-rg}.  Code written for this reimplementation.\\
    2020-005  \cite{cert-2020-005} & neuroscience & Larisch \\
    2020-006  \cite{cert-2020-006} & neuroscience & Detorakis \\
    2020-007  \cite{cert-2020-007} & neuroscience & \textbf{in progress} \\
    2020-008  \cite{cert-2020-008} & COVID-19 & Modelling of
    interventions on COVID-19 cases in the UK; checked as preprint (NOW REMOVED XXX)
    and now published \cite{Davies2020-vj} \\
    2020-009  \cite{cert-2020-009} & COVID-19 & tracing \\
    2020-010  \cite{cert-2020-010} & COVID-19 & report 9 \\
    2020-011  \cite{cert-2020-011} & COVID-19 & Modelling of Covid
    spread across Europe;  paper was checked when in press (now  published) \cite{Flaxman2020-yb}. \\
    2020-012  \cite{cert-2020-012} & COVID-19 & Modelling of COVID-19 spread across the USA; preprint currently review \cite{Unwin2020}. \\
    2020-013  \cite{cert-2020-013} & neuroscience & biorXiv \cite{Spitschan2020.06.02.129502}. \\
    2020-014  \cite{cert-2020-014} & neuroscience & eLife \cite{Sadeh2020}. \\
    2020-015  \cite{cert-2020-015} & neuroscience & eLife \cite{Liou2020}. \\
    2020-016  \cite{cert-2020-016} & GIScience & 
    R models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020}. \\
    2020-017  \cite{cert-2020-017} & GIScience & 
    Spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020}. \\
    2020-018  \cite{cert-2020-018} & GIScience & 
    Cellular automaton in R for modeling dynamic phenomena within a Discrete Global Grid System (DGGS) data model; using a subset for demonstration \cite{Hojati2020}. \\
    2020-019  \cite{cert-2020-019} & GIScience & Reachability analysis of suburban transportation using analysis of privately owned shared cars; workflow uses Python, and reproduction with a considerably subsampled dataset \cite{Illium2020}. \\
    2020-020  \cite{cert-2020-020} & GIScience & In-database windows operators for processing spatio-temporal data streams based on Python and PostgreSQL/PostGIS \cite{Werner2020}; reproduction using container. \\
    2020-021  \cite{cert-2020-021} & GIScience & Python code for comparing supervised machine learning models for spatial nominal entity recognition. \cite{Medad2020}. \\
    2020-022  \cite{cert-2020-022} & GIScience & Python code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}. \\
    2020-023  \cite{cert-2020-023} & GIScience & Analysis of spatial footprints of geo-tagged extreme weather events from social media using R \cite{Owuor2020}. \\
    \\ \bottomrule
  \end{tabular}
  \caption{Register of completed certificates as of October 2020.  An interactive version
  is available at \url{http://codecheck.org.uk/register}. % TODO: SJE to complete...
  }
  \label{tab:register}
\end{table}

@@

To demonstrate the typical contents of a certificate, in
Figure~\ref{fig:cert} we have annotated the first four out of ten
pages of the certificate \texttt{2020-012} \cite{cert-2020-012}.
The referenced article \cite{Unwin2020} describes the modelling COVID-19
across the USA.
%% TODO - has this been published yet?.  No, not yet... in review...
Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
which points to the certificate as archived on Zenodo.
The CODECHECK logo is
attached to the certificate to denote successful reproduction.
Figure~\ref{fig:cert}B provides the key metadata about the check --
which paper was checked, who the authors and the codechecker were,
and when it was checked. Materials
relating to the codecheck are freely available via the Repository
link. This table is generated from machine readable text file, the
CODECHECK configuration file \texttt{codecheck.yml}.
The configuration file collects the core pieces of metadata surrounding
a check and enables current and future automation of workflows and 
meta-analyses.
Figure~\ref{fig:cert}C is a textual summary of how the codecheck was
performed, and any interesting findings.

Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
lists the outputs that were generated from the codecheck (this is
termed the MANIFEST).  For each output, there is a comment stating
which figure/table it should be compared to in the original paper,
along with the size of the file.  On page 3 of the certificate,
Figure~\ref{fig:cert}E shows some more detailed notes from the
codechecker, in this case indicating what steps were needed to
initialize the environment and run the code.  This particular example
was quite long, taking about 17 hours to run.  Finally, on page 4 of
the certificate, we see the first output that was generated by the
codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
figure~4 of \cite{Unwin2020}.  Subsequent pages of the certificate show other
outputs (not shown here).

>>>>

We have a curated list of 22 certificates available at
\url{https://codecheck.org.uk/register}. These are a mixture of
reproductions of papers for a variety of reasons and fall into three themes:
(1) `classic' papers from computational neuroscience
(2) COVID-19 modelling preprints, and
(3) GIScience.  
The first theme was the initial set of papers used to explore the concept
of CODECHECK. The idea was to take very well known historical articles from a 
specific domain, in this case the work area of co-author SJE, possibly 
as as a conversation starter for potential collaborating journals.
The second theme is rather opportunistic. As the COVID-19 pandemic swept
over the whole Earth, we hoped to contribute a small part to science's answer
to the unknown virus. The checks were solicited through community interaction
%Several COVID-19 models have been done during the time of peer review of these papers,
and under our initiative rather than requested from journal, but the certificate
has then been acknowledged/cited in the paper \cite{Davies2020-vj}.
The third theme represents co-author DN's service as Reproducibility
Reviewer at the AGILE conference series.
Independent of CODECHECK, the \emph{Reproducible AGILE} Initiative \cite{reproducible_agile}
established a process for reproducing workflows at the AGILE conference
series \cite{nust_improving_2020}.
While using slightly different terms and infrastructure, e.g., ``reproducibility
reports'' are published on OSF instead of certificates on Zenodo, the 
reproducibility reviews adhere to the codecheck principles.
Furthermore, a small number of checks was completed as part of a peer reviews
for GIScience journals.

@@

\begin{table}
  \footnotesize

  \centering

  \begin{tabular}{llp{12cm}}
    \toprule
    \textbf{Certificate} & \textbf{Research area} & \textbf{Description} \\ \midrule
    2020-001 \cite{cert-2020-001} & Machine learning & Code for benchmarking ML classification tool checked post acceptance of manuscript and before its publication in \textit{Gigascience} \cite{Piccolo2020-lo}. \\
    2020-002  \cite{cert-2020-002} & Neuroscience & Code written for this project checked by second project member as demonstration using paper from 1997 showing unsupervised learning from natural images \cite{Hancock1992-mp}. \\
    2020-003  \cite{cert-2020-003} & Neuroscience &  Code written for this project checked by second project member as demonstration using classic paper on models of associative memory \cite{Hopfield1982-mz}. \\
    2020-004  \cite{cert-2020-004} & Neuroscience & Code written for this project checked by second project member as demonstration using classic paper on cart-pole balancing problem \cite{Barto1983-rg}. \\
    2020-005  \cite{cert-2020-005} & Neuroscience & Check of independent reimplementation of spike-timing-dependent plasticity (STDP) model \cite{larisch_re_2019} conducted as demonstration for this paper. \\ % Python ANNarchy
    2020-006  \cite{cert-2020-006} & Neuroscience & Check of independent reimplementation of a generalized linear integrate-and-fire neural model \cite{detorakis_re_2017} conducted as demonstration for this paper. \\
    2020-007  \cite{cert-2020-007} & Neuroscience & Check of independent reimplementation of analysing spike patterns of neurons \cite{hathway_re_2018} conducted as demonstration for this paper. \\
    2020-008  \cite{cert-2020-008} & COVID-19 & Code for modelling of interventions on COVID-19 cases in the UK checked at preprint stage \cite{davies-preprint-2020} and later published \cite{Davies2020-vj}. \\
    2020-009  \cite{cert-2020-009} & COVID-19 & Code for analysis of effectiveness of measures to reduce transmission of SARS-CoV-2 checked as preprint \cite{kucharski-preprint-2020} and later published \cite{kucharski_effectiveness_2020}. \\ % R, check mentiond in acks, but not linked
    2020-010  \cite{cert-2020-010} & COVID-19 & Code for analysis of non-pharmaceutical interventions (Report 9) checked as a preprint \cite{ferguson_report_2020}. \\ % CovidSim
    2020-011  \cite{cert-2020-011} & COVID-19 & Code for modelling of COVID-19 spread across Europe was provided by authors and checked while paper was in press \cite{flaxman_estimating_2020}. \\
    2020-012  \cite{cert-2020-012} & COVID-19 & Code for modelling of COVID-19 spread across the USA was checked as preprint \cite{unwin_report_2020} and later published \cite{unwin_state-level_2020}. \\
    2020-013  \cite{cert-2020-013} & Neuroscience & Code for analsis of rest-activity patterns in people wihout con-mediated vision was checked as a preprint \cite{Spitschan2020.06.02.129502} after direct contact with the authors. \\ % MATLAB
    2020-014  \cite{cert-2020-014} & Neuroscience & Code for analysis of pertubation patterns of neural activity was checked after publication as part of publisher collaboration \cite{Sadeh2020}. \\ % MATLAB
    2020-015  \cite{cert-2020-015} & Neuroscience & Code for a neural network model for human focal seizures was checked after publication as part of publisher collaboration \cite{Liou2020}. \\ % MATLAB
    2020-016  \cite{cert-2020-016} & GIScience & Code for models demonstrating the Modifiable Aral Unit Problem (MAUP) in spatial data science \cite{Brunsdon2020} was checked during peer review. \\ % R
    2020-017  \cite{cert-2020-017} & GIScience & Code for spatial data handling, analysis, and visualisation using a variety of R packages \cite{Bivand2020} was checked after peer review before publication. \\ % R
    2020-018  \cite{cert-2020-018} & GIScience & AGILE conference reproducibility report using a demonstration data subset with cellular automaton for modeling dynamic phenomena \cite{Hojati2020}. \\ % R
    2020-019  \cite{cert-2020-019} & GIScience & AGILE conference reproducibility report with subsampled dataset for reachability analysis of suburban transportation using shared cars \cite{Illium2020}. \\
    2020-020  \cite{cert-2020-020} & GIScience & AGILE conference reproducibility report using a container for checking in-database windows operators for processing spatio-temporal data \cite{Werner2020}. \\ % Python + PostGIS
    2020-021  \cite{cert-2020-021} & GIScience & AGILE conference reproducibility report checking code for comparing supervised machine learning models for spatial nominal entity recognition \cite{Medad2020}. \\ % Python
    2020-022  \cite{cert-2020-022} & GIScience & AGILE conference reproducibility report checking code for visualising text analysis on intents and concepts from geo-analytic questions \cite{Xu2020}. \\ % Python
    2020-023  \cite{cert-2020-023} & GIScience & AGILE conference reproducibility report on analysis of spatial footprints of geo-tagged extreme weather events from social media \cite{Owuor2020}. \\
    2020-024  \cite{cert-2020-024} & Neuroscience & Code for multi-agent system for concept drift detection in electromyography \cite{vieira_driftage_2020} was checked during peer review. \\ % R
    2020-025  \cite{cert-2020-025} & GIScience & Adaptation and application of Local Indicators for Categorical Data (LICD) to archaeological data \cite{carrer_application_2021} was checked after peer review before publication. \\ % R
    \\ \bottomrule
  \end{tabular}
  \caption{Register of completed certificates as of October 2020.  An interactive version
  is available at \url{http://codecheck.org.uk/register}. % TODO: SJE to complete...
  }
  \label{tab:register}
\end{table}

@@

To demonstrate the typical contents of a certificate, in
Figure~\ref{fig:cert} we have annotated the first four out of ten
pages of the certificate \texttt{2020-012} \cite{cert-2020-012}.
The referenced article \cite{unwin_report_2020} describes the modelling COVID-19
across the USA.
%% TODO - has this been published yet?.  No, not yet... in review...
Figure~\ref{fig:cert}A shows the certificate number, and its DOI,
which points to the certificate as archived on Zenodo.
The CODECHECK logo is
attached to the certificate to denote successful reproduction.
Figure~\ref{fig:cert}B provides the key metadata about the check --
which paper was checked, who the authors and the codechecker were,
and when it was checked. Materials
relating to the codecheck are freely available via the Repository
link. This table is generated from machine readable text file, the
CODECHECK configuration file \texttt{codecheck.yml}.
The configuration file collects the core pieces of metadata surrounding
a check and enables current and future automation of workflows and 
meta-analyses.
Figure~\ref{fig:cert}C is a textual summary of how the codecheck was
performed, and any interesting findings.

Figure~\ref{fig:cert}D (page 2 of the certificate) shows a table that
lists the outputs that were generated from the codecheck (this is
termed the MANIFEST).  For each output, there is a comment stating
which figure/table it should be compared to in the original paper,
along with the size of the file.  On page 3 of the certificate,
Figure~\ref{fig:cert}E shows some more detailed notes from the
codechecker, in this case indicating what steps were needed to
initialize the environment and run the code.  This particular example
was quite long, taking about 17 hours to run.  Finally, on page 4 of
the certificate, we see the first output that was generated by the
codecheck (Figure~\ref{fig:cert}F).  In this case, the figure matches
figure~4 of \cite{unwin_report_2020}.
Subsequent pages of the certificate show other outputs (not shown here).
